What defines big team science?

### 2025-06-03 manuscript_definition.Rmd
updating figures, references, and discussionn

    Significance statement: This work is the first to examine big team science authorship (i.e., 10+ authors) across millions of published works. Big teams can provide high-impact, important research within scientific publishing, and this report suggests a promising trend of increasing numbers of publications that increasingly represent earlier career and varied scholars. The number of geopolitical entities for researcher affiliation is increasing over time, showing the results of globalization and the ability to connect across time zones and cultures. While publications are generally diversifying, we did not yet find equality in the representation for first or corresponding authors. First authors appear to be less diverse, representing more European and North American authors, while other authors include more Asian and African authors.

@@

  Pre-registration: This manuscript was preregistered with the same conceptual ideas using Google Scholar and ORC-ID databases (https://osf.io/f2dtr) but then was updated with access to the Scopus database for a broader picture of BTS projects (https://osf.io/fheun).

@@

keywords          : "big team, science, authorship, credit"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
classoption       : "man"
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: inline
bibliography: references.bib
appendix          :
  - "appendix.Rmd"

@@

# things to import
var_list <- c("author_stats", "paper_stats", "author_paper_year",
              "career_time", "lm_coefficients", "geo_country")

@@

The movement toward team science reflects demands of modern research to
answer increasingly complex questions, meet funder and university desires
for interdisciplinary research, and the desire to increase diversity of
scientific representation [@council2015]. Further, the evolution of team
science reflects broader shifts in research practices, driven by two
sources: 1) increasing globalization and technology that allows for
real-time interdisciplinary research [@jones2008], and 2) expanding
interest in reproducibility, replication, and generalizability [i.e.,
the credibility movement, @maxwell2015; @nelson2018; @zwaan2018;
@vazire2022]. Technological advances have provided easier ways to
collaborate with people who are from other universities and countries
through document sharing platforms (e.g., Google, GitHub, and the Open
Science Framework), video chatting platforms (e.g., Zoom, Microsoft
Teams), and messaging and project management platforms (e.g., Slack,
Trello, when2meet, etc.).

The credibility movement has been the main catalyst for the increased
interest and broader shift of research practices. The movement
emphasizes reproducibility and transparency in science, encouraging
researchers to form new ways to increase the rigor in scientific
endeavors. Throughout the last decade, the credibility movement has pushed for larger, more diverse teams and the involvement of participants from varied backgrounds. This shift in teams and participants focuses on increasing credibility, generalizability, and reliability of scientific findings. This form of collaboration has been coined “Big Team Science.”
Big Team Science (BTS) builds on team science by scaling efforts to
include larger, often globally diverse teams, which requires significant
coordination and infrastructure [@coles2022; @forscher2022a;
@stewart2017]. BTS projects and organizations organize extensive
collaborations, intentionally incorporating diverse populations and
perspectives into research. This large-scale approach enhances the
reliability and generalizability of findings by integrating varied
methodologies and viewpoints, leading to more robust and inclusive
scientific outcomes. BTS organizations often pool extensive networks of
researchers and resources, aiming to tackle grand scientific challenges
that would be difficult to address within smaller or less coordinated
collaborations. By having both collaborations that span across the globe
and subfields of research areas, age groups, and education levels should
help to drive science in the path of better materials, reliability,
generalizability, and more robust sample sizes (when necessary) in a
study [@auspurg2021; @nosek2014method; @lebel2018].

@@

Beyond replication concerns, the credibility movement has mirrored calls
for diversification or de-WEIRDing (e.g., Western, Educated,
Industrialized, Rich, and Democratic) scientific research [@henrich2010;
@newson2021; @rad2018] by improving representation in research samples.
Like the large-scale studies in Physics [@aphilos2021; @castelnovo2018]
and Biology [@collins2003], the Social Sciences struggle to represent
the breadth of humanity across both researcher and population
characteristics. Now, grassroots organizations, such as the
Psychological Science Accelerator [@moshontz2018], ManyBabies
(<https://manybabies.github.io/>), NutNet (<https://nutnet.org/>), and
DRAGNet (<https://dragnetglobal.weebly.com/>) can begin to tackle these
issues by recruiting research labs from all over the globe to provide
diversity in geographic, linguistic, and researcher representation.
Publications have examined the global understanding of morality, face
processing, COVID-19 information signaling, and more [@bago2022;
@dorison2022; @jones2021;
@psychologicalscienceacceleratorself-determinationtheorycollaboration2022;
@vanbavel2022; @wang2021]. While these organizations and one-time groups
for BTS studies have provided an incredible wealth of data for the
scientific community, we do not yet know exactly how to define BTS -
which is generally termed "an unusually large number of collaborators"
[@forscher2022a; @coles2022].

The lack of formal definition raises questions about whether it
represents a distinct phenomenon or simply a natural extension of team
science. These big teams pose unique challenges, including coordinating
work across diverse time zones, managing conflicts in decision-making,
and ensuring fair distribution of credit for contributions
[@cummings2007; @wuchty2007], but also could provide big rewards by
pooling expertise and increased interdisciplinary funding [@fiore2008].
This paper seeks to clarify the concept of BTS by first establishing a
data-driven definition based on publication patterns. With this
quantitative distinction in place, we analyze publication trends over
time to assess the trajectories of both traditional team science and
BTS. Furthermore, we investigate the diversity of authors involved in
these collaborations to explore whether shifts in the scientific
landscape—such as efforts to de-WEIRD science and the expansion of
collaborative opportunities—have influenced who participates in team
science and BTS. By synthesizing insights from the growth and
diversification of team science, this paper seeks to critically examine
the emergence of big teams. Specifically, it aims to explore whether big
teams are quantitatively different from traditional collaboration models
with the following research questions.

@@

-   Research Question 1: Exploring historical and current publication
    values, what should define big team science versus team science?
    -   Question 1A: What number of authors and institutional
        affiliations should designate the differences between team
        science and big team science?
    -   Question 1B: Using the definition from 1A, are there changes in
        the number of publications over time?
-   Research Question 2: How has the diversity of those involved in big
    team science versus team science changed over time?

@@

We used data from 1970-2024 in the Scopus database, as it is noted
online that 1970 and forward includes cited references for calculation
of several of our variables described below. We analyzed our results
based on four subject areas present in the Scopus database: Physical
Sciences, Health Sciences, Social Sciences, and Life Sciences. We
filtered the database to include articles, articles in press, business
articles, conference papers, data papers, preprints, and surveys using
Elsevier's classification system. This project was supported by access
to the Scopus database through the International Center for the Study of
Research.

@@

### RQ2: Author Diversity Statistics

#### Seniority

@@

#### Geopolitical Region

@@

The total number of papers included in the Scopus database at the time
of this analysis was `r total_papers`. `r total_papers_past1970`
articles were included past 1970 in the defined article types, which
included `r distinct_authors_past1970` distinct authors. We then
filtered the data to include only teams, which was defined as two
authors from at least two institutions. The total number of papers for
team projects was `r team_author_aff_papers` and
`r team_author_aff_author` distinct authors. The data was then
classified into subject areas by paper, which lead to missing data. The
final number of papers included was `r team_author_aff_papers_noNA` with
`r team_author_aff_author_noNA` distinct authors. The dataset was
curated to include one row per author, paper, and subject area (i.e.,
long format [@wickham2007]) which included
`r team_author_aff_subject_noNA` total rows of data. For analyses and
descriptive statistics below, we include sample sizes for clarity.

```{r stats-authors-fig, include = F, fig.cap="The left panel depicts the number of authors included on a paper by subject area, and the right panel demonstrates the number of affiliations by subject area. The boxplot shows the median (bold line), the interquartile range (the box), and the minimum to the 90th percentile of the number of authors/affiliations as the range line. Normally these plots include the entire range of the data, but these extreme range made the boxplot information unreadable. The dots indicate the average number of authors/affiliations for each area with the size of the dot indicating the standard deviation of the statistic. Therefore, larger dots indicate more variability in the number of authors and affiliations."}
temp_graph <- DF$author_stats %>% 
  filter(SubjectArea != "Multidisciplinary")

@@

auth_plot | aff_plot
```

\@ref(fig:stats-authors-fig) displays the number of authors and
affiliations by subject area. The figure demonstrates that the median
number of authors is largest for health sciences, followed by life
science, physical sciences, and then social sciences. The general
pattern of team authorship includes about 2-8 authors, from about 2-4
institutions. We used the maximum (i.e., across all subject areas) 90th
percentile as our exploratory definition for big team papers after
examining the results from this analysis. We selected this percentile to
have the high of the distribution, but also to be able to include enough
papers for analysis across time. Therefore, big teams were defined as 11
authors from at least 6 different institutions[\^In a previous version
of this manuscript, we defined big teams as 10 authors from at least 10
institutions based on our own experiences working with research
consortium. All definitions are likely subjective, but the definition in
this manuscript represents the top 10% of author and affiliations in a
large body of papers].

\@ref(tab:big-teams-table) includes the number of distinct authors and
papers for each subject area by overall teams and big teams using our
90th percentile definition. The total number of distinct authors for big
team papers was `r big_teams_papers` with `r big_teams_authors` distinct
authors. In RQ2, we split the big team data into small (11-49), medium
(50-99), and large big team (100+) groupings for convenience to
display/analyze geopolitical regions. The table shows the number of
authors and papers for those analyses.

@@

SEE IF YOU CAN GET ALL THE NUMBERS SOME ARE MISSING

For analyzing changes across time, we split the data into team science
projects (2-10 authors, 2-5 affiliations) and BTS projects (as defined
above, 11+ authors, 6+ affiliations). The number of papers found in
Scopus across time for each subject area are displayed in Figure
\@ref(paper-time-fig). The visual results indicated that the number of
team science papers was increasing the most in physical sciences for all
manuscripts, followed by life and health sciences, and the last is
social sciences. Examining only BTS projects shows that the rate is also
increasing across time. All teams appear to start increasing in the
1990s, while BTS projects do not start increasing off floor effects
until past 2000. The health and life sciences show the largest increases
across time in big teams with the smallest trend in the social sciences.

```{r paper-time-fig, include = F, fig.cap = "The number of manuscripts across time for all team science papers (left) and big team science papers (right)."}
temp <- DF$author_paper_year %>% 
  filter(SubjectArea != "Multidisciplinary") %>% 
  filter(Year < 2024) %>% 
  mutate(SubjectArea = gsub("Sciences", "", SubjectArea))

p1 <- ggplot(temp %>% 
         filter(team_size == "regular"), 
         aes(x = Year, y = Number_Papers, 
             colour = SubjectArea)) + 
  geom_point() + 
  theme_minimal() + 
  scale_color_discrete(name = "Subject Area") + 
  theme(legend.position = "bottom",
        axis.text.y = element_text(angle = 45, hjust = 1)) + # Angles the y-axis text
  ylab("Number of Papers") + 
  ggtitle("Team Science Projects")

p2 <- ggplot(temp %>% 
         filter(team_size == "big"), 
         aes(x = Year, y = Number_Papers, 
             colour = SubjectArea)) + 
  geom_point() + 
  theme_minimal() + 
  scale_color_discrete(name = "Subject Area") + 
  theme(legend.position = "bottom",
        axis.text.y = element_text(angle = 45, hjust = 1)) + # Angles the y-axis text
  ylab(NULL) +  # Removes the y-axis label
  ggtitle("BTS Projects")

@@

combined_plot
```

```{r rate-estimates-fig, include = F, fig.cap = "Exponential growth rate estimates with 95% confidence intervals."}
# Initialize the results list
corr.results <- list()

@@

corr.DF <- bind_rows(corr.results) %>% 
  mutate(group = rep(group, each = 8)) %>% 
  mutate(SubjectArea = c(
    rep(subjarea, each = 2),
    rep(subjarea, each = 2)
    )) %>% 
  filter(term != "a") %>% 
  mutate(conf.low = estimate-2*std.error,
         conf.high = estimate+2*std.error)

ggplot(corr.DF, aes(x = SubjectArea, y = estimate, color = group)) + 
  geom_point() +
  theme_minimal() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = .2) + 
  ylab("Rate Estimate") + 
  xlab("Subject Area") + 
  scale_color_discrete(name = "Team Size", 
                       labels = c("BTS", "Team Science")) + 
  theme(legend.position = "bottom") + 
  geom_hline(yintercept = 0)
```

Using the `minpack.lm` library [@elzhov2023], we calculated the
exponential rate of growth for team science and BTS projects, and these
results are shown in \@ref(fig:rate-estimates-fig). All growth rate
confidence intervals excluded zero, indicating an exponential increase
in the number of team papers over time. BTS growth rates were
always higher than their team science counterparts, but the 95%
confidence intervals for the growth estimate overlapped for all
statistics. Therefore, the growth trends, while visually appearing to be
different, were likely similar for each subject area and team size when
examined by estimating exponential growth statistics.

## RQ2: Diversity Statistics

### Seniority

```{r fig-career, include = TRUE,  warning = FALSE, message = FALSE, fig.cap = "Average career length for big-team science authors. Larger dots indicate more variability in career length for authors by averaging the standard deviation in career length for each manuscript within a year. The data has been filtered to at least 10 publications in a year for this graph.", fig.width=10, fig.height=6}
ggplot(DF$career_time %>% 
         filter(SubjectArea != "Multidisciplinary") %>% 
         filter(num_papers >= 10) %>% 
         filter(Year < 2024) %>% 
         mutate(Team = factor(Team, 
                              levels = c("Regular", "Big"),
                              labels = c("Team Science", "BTS"))), 
       aes(Year, avg_career, color = Team)) +
  theme_classic(base_size = 15) + 
  geom_point(aes(size = sd_career)) +
  facet_wrap(~SubjectArea) + 
  theme(legend.position = "bottom") + 
  ylab("Average Career Length") +
  xlab("Year of Publication") + 
  scale_color_manual(values = c("#D55E00", "#0072B2"), name = "Team Size") + # Red and Blue from RdBu ends
  guides(size = "none")
  #scale_size_continuous(name = "Number Papers")

@@

@@

# create upper and lower CIs for each coefficient
DF$lm_coefficients <- DF$lm_coefficients %>% 
  mutate(b_avg_low = b_avg_effect - qt(.025, sample_size, lower.tail = F)*se_avg_effect,
         b_avg_high = b_avg_effect + qt(.025, sample_size, lower.tail = F)*se_avg_effect,
         b_std_low = b_std_effect - qt(.025, sample_size, lower.tail = F)*se_std_effect,
         b_std_high = b_std_effect - qt(.025, sample_size, lower.tail = F)*se_std_effect)

@@

```{r fig-heatmap, include = TRUE, fig.cap="Heatmap results of regression analyses for career length, number of publications, and geopoliticalwithin the region. Each square represents a *b* value or the slope of the predictor (x-axis) onto the dependent variable (each panel), with the exception of the bottom row which is the effect size of each regression analysis $R^2$. Slopes included both the overall value of the predictor ($b$, $b_M$) and the standard deviation of the predictor over time ($b_{SD}$). The color of the square represents the strength of the predictor. The top figure represents all results together for comparison across analyses. The bottom row represents individual heatmaps for each hypothesis to distinguish small differences between subject areas for those research questions. Non-significant results are indicated with NS on the plot.", warning = FALSE, message = FALSE, fig.width=10, fig.height=6}

graphDF <- DF$lm_coefficients %>% 
  select(Discipline, b_avg_effect, b_std_effect, type, rsquared, coefficient) %>% 
  pivot_longer(cols = c(b_avg_effect, b_std_effect, rsquared)) %>% 
    left_join(
    DF$lm_coefficients %>% 
  select(Discipline, type, coefficient, r2_high:b_std_high) %>% 
  pivot_longer(cols = c(-Discipline, -type, -coefficient)) %>% 
  mutate(ci_name = ifelse(
    grepl("low", name), "low", "high"
  ),
  name = gsub("r2_low|r2_high", "rsquared", name),
  name = gsub("b_avg_low|b_avg_high", "b_avg_effect", name),
  name = gsub("b_std_low|b_std_high", "b_std_effect", name)),
  by = c("Discipline", "type", "coefficient", "name")
  ) %>% 
  mutate(name = ifelse(
    coefficient == "diversity" & name != "rsquared", "b_effect", name
  )) %>% 
  mutate(name = factor(name, 
                levels = c("b_avg_effect", "b_effect", 
                           "b_std_effect", "rsquared"),
                labels = c("M Coefficient", "Coefficient", 
                           "SD Coefficient", "Effect Size"))) %>%
  rename(Estimate = value.x,
         Regression = coefficient,
         CI = value.y,
         Size = type,
         Coefficient = name) %>% 
  mutate(Discipline = gsub(" Sciences", "", Discipline)) %>% 
  pivot_wider(
    id_cols = c(Discipline, Size, Regression, Coefficient, Estimate),
    names_from = ci_name,
    values_from = CI
  ) %>%
  mutate(heat_label = ifelse(
    low < .00001 & high > .00001, "NS", ""
  )) %>% 
  mutate(Size = factor(Size, 
                       levels = c("regular", "big"),
                       labels = c("Regular", "Big")))

@@

updated binning:

-   teams: 2-10 authors, 2-5 inst
    -   first and last author only versus everyone else
-   big ass teams: 6+ affs
    -   small: 11-49
        -   first three plus last versus everyone else
    -   medium: 50-99
        -   first five plus last versus everyone else
    -   large: 100+
        -   first 10 plus last versus everyone else

for time: teams versus big teams combining results separate panel of the
big teams separated (not by time)

```{r small-effects, warning = FALSE}
DF$small_first_gp <- DF$small_first_gp %>% 
  filter(aff_un != "other") %>% 
  pivot_wider(id_cols = Year, 
              names_from = aff_un, 
              values_from = count,
              values_fill = 0)

>>>>

    Significance statement: This study is the first large-scale, data-driven analysis of authorship in Big Team Science (BTS), defined here as collaborations with 11 or more authors across six or more institutions. Drawing on millions of published works across four scientific domains, we show that BTS has grown rapidly in recent years and increasingly includes early-career researchers and contributors from a broader range of geopolitical regions. However, inclusion in BTS is not equally distributed. While author lists are becoming more globally diverse overall, leadership roles—particularly first and corresponding authorships, remain disproportionately concentrated in high-income countries, especially in Europe and North America. These findings highlight both the promise and limitations of BTS as a vehicle for global, inclusive science, underscoring the need for structural reforms to ensure equity in recognition and participation.

@@

  Pre-registration: This manuscript was preregistered with the same conceptual ideas using Google Scholar and ORC-ID databases (https://osf.io/f2dtr) but then was updated with access to the Scopus database for a broader picture of BTS projects (https://osf.io/fheun). After review, we updated the manuscript to reflect that the definition of big team science is unclear, and focused on a data-driven definition to explore our research questions.

@@

keywords          : "big team, science, authorship, credit"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
classoption       : "man"
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: inline
bibliography: references.bib

@@

# things to import
var_list <- c("author_stats", "paper_stats", "author_paper_year",
              "career_time", "lm_coefficients", "geo_country",
              "effect_size_gp")

@@

The movement toward team science reflects demands of modern research to
answer complex questions, meet funder and university desires for
interdisciplinary research, and the desire to increase diversity of
scientific representation [@council2015]. Further, the evolution of team
science reflects broader shifts in research practices, driven by two
sources: 1) increasing globalization and technology that allows for
real-time interdisciplinary research [@jones2008], and 2) expanding
interest in reproducibility, replication, and generalizability [i.e.,
the credibility movement, @maxwell2015; @nelson2018; @zwaan2018;
@vazire2022]. Technological advances have provided easier ways to
collaborate with people who are from other universities and countries
through document sharing platforms (e.g., Google, GitHub, and the Open
Science Framework), video chatting platforms (e.g., Zoom, Microsoft
Teams), and messaging and project management platforms (e.g., Slack,
Trello, when2meet, etc.).

The credibility movement has been the main catalyst for the increased
interest and broader shift of research practices. The movement
emphasizes reproducibility and transparency in science, encouraging
researchers to form new ways to increase the rigor in scientific
endeavors. Throughout the last decade, the credibility movement has
pushed for larger, more diverse teams and the involvement of
participants from varied backgrounds. This shift in teams and
participants focuses on increasing credibility, generalizability, and
reliability of scientific findings. This form of collaboration has been
coined "Big Team Science." Big Team Science (BTS) builds on team science
by scaling efforts to include larger, often globally diverse teams,
which requires significant coordination and infrastructure [@coles2022;
@forscher2022a; @stewart2017]. BTS projects and organizations organize
extensive collaborations, intentionally incorporating diverse
populations and perspectives into research. This large-scale approach
enhances the reliability and generalizability of findings by integrating
varied methodologies and viewpoints, leading to more robust and
inclusive scientific outcomes. BTS organizations often pool extensive
networks of researchers and resources, aiming to tackle grand scientific
challenges that would be difficult to address within smaller or less
coordinated collaborations. By having both collaborations that span
across the globe and subfields of research areas, age groups, and
education levels should help to drive science in the path of better
materials, reliability, generalizability, and more robust sample sizes
in a study [@auspurg2021; @nosek2014method; @lebel2018].

@@

Beyond replication concerns, the credibility movement has mirrored calls
for diversification or de-WEIRDing (e.g., Western, Educated,
Industrialized, Rich, and Democratic) scientific research [@henrich2010;
@newson2021; @rad2018] by improving representation in research samples.
Like the large-scale studies in Physics [@aphilos2021; @castelnovo2018]
and Biology [@collins2003], the Social Sciences struggle to represent
the breadth of humanity across both researcher and population
characteristics. Now, grassroots organizations, such as the
Psychological Science Accelerator [@moshontz2018], ManyBabies
(<https://manybabies.github.io/>), NutNet (<https://nutnet.org/>),
DRAGNet (<https://dragnetglobal.weebly.com/>), and IceCube
(<https://icecube.wisc.edu/>) can begin to tackle these issues by
recruiting research labs from all over the globe to provide diversity in
geographic, linguistic, and researcher representation. Publications have
examined the global understanding of morality, face processing, COVID-19
information signaling, and more [@bago2022; @dorison2022; @jones2021;
@psychologicalscienceacceleratorself-determinationtheorycollaboration2022;
@vanbavel2022; @wang2021]. While these organizations and one-time groups
for BTS studies have provided an incredible wealth of data for the
scientific community, we do not yet know exactly how to define BTS -
which is generally termed "an unusually large number of collaborators"
[@forscher2022a; @coles2022].

The lack of formal definition raises questions about whether it
represents a distinct phenomenon or simply a natural extension of team
science. These big teams pose unique challenges, including coordinating
work across diverse time zones, managing conflicts in decision-making,
and ensuring fair distribution of credit for contributions
[@cummings2007; @wuchty2007], but also could provide big rewards by
pooling expertise and increased interdisciplinary funding [@fiore2008].
This paper seeks to clarify the concept of BTS by first establishing a
data-driven definition based on publication patterns. With this
quantitative distinction in place, we analyze publication trends over
time to assess the trajectories of both traditional team science and
BTS. Furthermore, we investigate the diversity of authors involved in
these collaborations to explore whether shifts in the scientific
landscape, such as efforts to de-WEIRD science and the expansion of
collaborative opportunities, have influenced who participates in team
science and BTS. By synthesizing insights from the growth and
diversification of team science, this paper seeks to critically examine
the emergence of big teams. Specifically, it aims to explore whether big
teams are quantitatively different from traditional collaboration models
with the following research questions.

@@

-   Research Question 1: Exploring historical and current publication
    values, what should define big team science versus team science?
    -   Question 1A: What number of authors and institutional
        affiliations should designate the differences between team
        science and big team science?
    -   Question 1B: Using the definition from 1A, are there changes in
        the number of publications over time?
-   Research Question 2: How has the diversity of those involved in team
    science and big team science changed over time?

@@

We used data from 1970-2024 in the Scopus database, as it is noted
online that 1970 and forward includes cited references for calculation
of several of our variables described below. We analyzed our results
based on four subject areas present in the Scopus database: Physical
Sciences, Health Sciences, Social Sciences, and Life Sciences. We used
the subject area split to ensure one field did not dominate BTS
definitions and determine differences in trends across sub-areas of
science. We filtered the database to include articles, articles in
press, business articles, conference papers, data papers, preprints, and
surveys using Elsevier's classification system. This project was
supported by access to the Scopus database through the International
Center for the Study of Research.

@@

### RQ2: Seniority

@@

### RQ2: Geopolitical Region

@@

The total number of papers included in the Scopus database at the time
of this analysis was `r total_papers`. `r total_papers_past1970`
articles were included past 1970 in the defined article types, which
included `r distinct_authors_past1970` distinct authors. We then
filtered the data to include only teams, which was defined as two
authors from at least two institutions. The total number of papers for
team projects was `r team_author_aff_papers` and
`r team_author_aff_author` distinct authors. The data was then
classified into subject areas by paper, which lead to missing data. The
final number of papers included was `r team_author_aff_papers_noNA` with
`r team_author_aff_author_noNA` distinct authors. The dataset was
curated to include one row per author, paper, and subject area (i.e.,
long format [@wickham2007]) which included
`r team_author_aff_subject_noNA` total rows of data.

```{r fig-stats-authors, include = TRUE, fig.cap="The left panel depicts the number of authors included on a paper by subject area, and the right panel demonstrates the number of affiliations by subject area. The boxplot shows the median (bold line), the interquartile range (the box), and the minimum to the 90th percentile of the number of authors/affiliations as the range line. Normally these plots include the entire range of the data, but these extreme range made the boxplot information unreadable. The dots indicate the average number of authors/affiliations for each area with the size of the dot indicating the standard deviation of the statistic. Therefore, larger dots indicate more variability in the number of authors and affiliations.", fig.width = 10, fig.height = 6}
temp_graph <- DF$author_stats %>% 
  filter(SubjectArea != "Multidisciplinary")

@@

auth_plot | aff_plot

ggsave(filename = "figure/figure_1_number.png", width = 10, height = 6)
```

Figure \@ref(fig:fig-stats-authors) displays the number of authors and
affiliations by subject area. The figure demonstrates that the median
number of authors is largest for health sciences, followed by life
science, physical sciences, and then social sciences. The general
pattern of team authorship includes about 2-8 authors, from about 2-4
institutions. We used the maximum value (i.e., across all subject areas)
for the 90th percentile as our exploratory definition for big team
papers after examining the results from this analysis. We selected this
percentile to have the high of the distribution, but also to be able to
include enough papers for analysis across time. Therefore, big teams
were defined as 11 authors from at least 6 different institutions.[^1]

[^1]: In a previous version of this manuscript, we defined big teams as
    10 authors from at least 10 institutions based on our own
    experiences working within a research consortium. All definitions
    are likely subjective, but the definition in this manuscript
    represents the top 10% of author and affiliations in a large body of
    papers.

```{r echo = FALSE}
big_teams_papers <- 968765
big_teams_authors <- 4541369
```

Supplemental Table \@ref(tab:big-teams-table) includes the number of
distinct authors and papers for each subject area by overall teams and
big teams using our 90th percentile definition. The total number of
distinct authors for big team papers was `r big_teams_papers` with
`r big_teams_authors` distinct authors. In RQ2, we split the big team
data into small (11-49), medium (50-99), and large big team (100+)
groupings for convenience to display/analyze geopolitical regions. The
table shows the number of authors and papers for those analyses.

@@

For analyzing changes across time, we split the data into team science
projects (2-10 authors, 2-5 affiliations) and BTS projects (as defined
above, 11+ authors, 6+ affiliations). The number of papers found in
Scopus across time for each subject area are displayed in Figure
\@ref(fig:fig-paper-time). The visual results indicated that the number
of team science papers was increasing the most in physical sciences for
all manuscripts, followed by life and health sciences, and the last is
social sciences. Examining only BTS projects shows that the rate is also
increasing across time. All teams appear to start increasing in the
1990s, while BTS projects do not start increasing off floor effects
until past 2000. The health and life sciences show the largest increases
across time in big teams with the smallest trend in the social sciences.

```{r fig-paper-time, include = TRUE, fig.cap = "The number of manuscripts across time for all team science papers (left) and big team science papers (right).", fig.width = 10, fig.height = 6}
temp <- DF$author_paper_year %>% 
  filter(SubjectArea != "Multidisciplinary") %>% 
  filter(Year < 2025) %>% 
  mutate(SubjectArea = gsub("Sciences", "", SubjectArea))

p1 <- ggplot(temp %>% 
         filter(team_size == "regular"), 
         aes(x = Year, y = Number_Papers, 
             colour = SubjectArea)) + 
  geom_point() + 
  theme_minimal() + 
  scale_color_discrete(name = "Subject Area") + 
  theme(legend.position = "bottom",
        axis.text.y = element_text(angle = 45, hjust = 1)) + # Angles the y-axis text
  ylab("Number of Papers") + 
  ggtitle("Team Science Projects") +
  theme(legend.title = element_text(size = 15),
    legend.text = element_text(size = 12))

p2 <- ggplot(temp %>% 
         filter(team_size == "big"), 
         aes(x = Year, y = Number_Papers, 
             colour = SubjectArea)) + 
  geom_point() + 
  theme_minimal() + 
  scale_color_discrete(name = "Subject Area") + 
  theme(legend.position = "bottom",
        axis.text.y = element_text(angle = 45, hjust = 1)) + # Angles the y-axis text
  ylab(NULL) +  # Removes the y-axis label
  ggtitle("BTS Projects") + 
  theme(legend.title = element_text(size = 15),
    legend.text = element_text(size = 15))

@@

combined_plot

ggsave(filename = "figure/figure_2_acrosstime.png", width = 10, height = 6)
```

Using the `minpack.lm` library [@elzhov2023], we calculated the
exponential rate of growth for team science and BTS projects, and these
results are shown in Supplemental Figure \@ref(fig:rate-estimates-fig). All growth rate
confidence intervals excluded zero, indicating an exponential increase
in the number of team papers over time. BTS growth rates were always
higher than their team science counterparts, but the 95% confidence
intervals for the growth estimate overlapped for all statistics.
Therefore, the growth trends, while visually appearing to be different,
were likely similar for each subject area and team size when examined by
estimating exponential growth statistics.

```{r echo = FALSE, include=FALSE}
# Initialize the results list
corr.results <- list()

@@

corr.DF <- bind_rows(corr.results) %>% 
  mutate(group = rep(group, each = 8)) %>% 
  mutate(SubjectArea = c(
    rep(subjarea, each = 2),
    rep(subjarea, each = 2)
    )) %>% 
  filter(term != "a") %>% 
  mutate(conf.low = estimate-2*std.error,
         conf.high = estimate+2*std.error) %>% 
  mutate(group = factor(group, 
                        levels = c("regular", "big"),
                        labels = c("Team Science", "BTS")))

ggplot(corr.DF, aes(x = SubjectArea, y = estimate, color = group)) + 
  geom_point() +
  theme_minimal() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = .2) + 
  ylab("Rate Estimate") + 
  xlab("Subject Area") + 
  scale_color_manual(values = c("#D55E00", "#0072B2"), name = "Team Size") + 
  theme(legend.position = "bottom") + 
  geom_hline(yintercept = 0)

ggsave(filename = "figure/figure_3_growth.png", width = 10, height = 6)
```

## RQ2: Seniority

```{r fig-career, include = TRUE,  warning = FALSE, message = FALSE, fig.cap = "Average career length for big-team science authors. Larger dots indicate more variability in career length for authors by averaging the standard deviation in career length for each manuscript within a year. The data has been filtered to at least 10 publications in a year for this graph.", fig.width=10, fig.height=6}
ggplot(DF$career_time %>% 
         filter(SubjectArea != "Multidisciplinary") %>% 
         filter(num_papers >= 10) %>% 
         filter(Year < 2025) %>% 
         mutate(Team = factor(Team, 
                              levels = c("Regular", "Big"),
                              labels = c("Team Science", "BTS"))), 
       aes(Year, avg_career, color = Team)) +
  theme_classic(base_size = 15) + 
  geom_point(aes(size = sd_career)) +
  facet_wrap(~SubjectArea) + 
  theme(legend.position = "bottom") + 
  ylab("Average Career Length") +
  xlab("Year of Publication") + 
  scale_color_manual(values = c("#D55E00", "#0072B2"), name = "Team Size") + # Red and Blue from RdBu ends
  guides(size = "none")
  #scale_size_continuous(name = "Number Papers")

@@

### Career Length

@@

# create upper and lower CIs for each coefficient
DF$lm_coefficients <- DF$lm_coefficients %>% 
  mutate(b_avg_low = b_avg_effect - qt(.025, sample_size, lower.tail = F)*se_avg_effect,
         b_avg_high = b_avg_effect + qt(.025, sample_size, lower.tail = F)*se_avg_effect,
         b_std_low = b_std_effect - qt(.025, sample_size, lower.tail = F)*se_std_effect,
         b_std_high = b_std_effect + qt(.025, sample_size, lower.tail = F)*se_std_effect)

@@

```{r fig-heatmap, include = TRUE, fig.cap="Heatmap results of regression analyses for career length, number of publications, and geopoliticalwithin the region. Each square represents a *b* value or the slope of the predictor (x-axis) onto the dependent variable (each panel), with the exception of the bottom row which is the effect size of each regression analysis $R^2$. Slopes included both the overall value of the predictor ($b$, $b_M$) and the standard deviation of the predictor over time ($b_{SD}$). The color of the square represents the strength of the predictor. The top figure represents all results together for comparison across analyses. The bottom row represents individual heatmaps for each hypothesis to distinguish small differences between subject areas for those research questions.", warning = FALSE, message = FALSE, fig.width=10, fig.height=6}

graphDF <- DF$lm_coefficients %>% 
  select(Discipline, b_avg_effect, b_std_effect, type, rsquared, coefficient) %>% 
  pivot_longer(cols = c(b_avg_effect, b_std_effect, rsquared)) %>% 
    left_join(
    DF$lm_coefficients %>% 
  select(Discipline, type, coefficient, r2_high:b_std_high) %>% 
  pivot_longer(cols = c(-Discipline, -type, -coefficient)) %>% 
  mutate(ci_name = ifelse(
    grepl("low", name), "low", "high"
  ),
  name = gsub("r2_low|r2_high", "rsquared", name),
  name = gsub("b_avg_low|b_avg_high", "b_avg_effect", name),
  name = gsub("b_std_low|b_std_high", "b_std_effect", name)),
  by = c("Discipline", "type", "coefficient", "name")
  ) %>% 
  mutate(name = ifelse(
    coefficient == "diversity" & name != "rsquared", "b_effect", name
  )) %>% 
  mutate(name = factor(name, 
                levels = c("b_avg_effect", "b_effect", 
                           "b_std_effect", "rsquared"),
                labels = c("M Coefficient", "Coefficient", 
                           "SD Coefficient", "Effect Size"))) %>%
  rename(Estimate = value.x,
         Regression = coefficient,
         CI = value.y,
         Size = type,
         Coefficient = name) %>% 
  mutate(Discipline = gsub(" Sciences", "", Discipline)) %>% 
  pivot_wider(
    id_cols = c(Discipline, Size, Regression, Coefficient, Estimate),
    names_from = ci_name,
    values_from = CI
  ) %>%
  mutate(heat_label = ifelse(
    low < .00001 & high > .00001, "NS", ""
  )) %>% 
  mutate(Size = factor(Size, 
                       levels = c("big", "regular"),
                       labels = c("BTS", "TS")))

@@

```{r xsmall-effects, warning = FALSE}
DF$effect_size_gp <- DF$effect_size_gp %>% 
  mutate(aff_un = gsub("Mirconesia", "Micronesia", aff_un)) %>% 
  filter(Year < 2025)

DF$effect_size_gp_xsmall_first <- DF$effect_size_gp %>% 
  filter(tolower(aff_un) != "other") %>% 
  filter(source_file == "xsmall_first") %>% 
  pivot_wider(id_cols = Year, 
              names_from = aff_un, 
              values_from = count,
              values_fill = 0)
